\LARGE \bf Mean Estimation from One-Bit Measurements

### 2019-06-19 onebit_full_V2.tex
Proof of normal approximation to SGD

%\documentclass[letterpaper]{IEEEtran}
\documentclass[letterpaper, 11pt]{IEEEtran}      % Use this line for a4 paper

@@

%\tikzexternalize[mode=list and make]
\usepackage{algorithmicx}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage{pstool}
\usepackage[latin1]{inputenc}
\usetikzlibrary{arrows,shapes}
\usepackage{xifthen}
\usepackage{epic}
%\usepackage{caption}
\usepackage{epstopdf}

@@

\newcommand{\mmse}{\mathsf{mmse}}
\newcommand{\unif}{\mathsf{unif}}
\newcommand{\card}{\mathrm{card}}
\newcommand{\ARE}{\mathsf{ARE}}
\newcommand{\supp}{\mathrm{supp} }
\renewcommand\vec[1]{\ensuremath\boldsymbol{#1}}
\newenvironment{proof}{\paragraph*{Proof}}{\hfill$\square$ \newline}
\newcommand{\sgn}{\mathrm{sgn} }
\newcommand{\Prob}{\mathbb{P} }
\newcommand{\argmax}{\mathrm{argmax}}
\newcommand*{\QEDA}{\hfill\ensuremath{\square}}
\newcommand{\CEO}{\mathsf{CEO}}

@@

%\node[above of = source3, node distance = 1cm] (dist) {$X_i \sim {\mathcal N} \left(\theta, \sigma^2 \right)$};
\node[below of = source2, node distance = 0.5cm] {$\vdots$};

@@

We give particular attention to the ARE of estimators with respect to an asymptotically normal efficient estimator that is not subject to the bit constraint. Specifically, let $\{a_n,\,n\in \mathbb N\}$ be a sequence such that 
\[
\sqrt{a_n}\left(\hat{\theta}_n - \theta\right) \overset{D}{\longrightarrow} \mathcal N(0, \sigma^2).
\]
Then the ARE of $\hat{\theta}_n$ with respect to an unconstrained efficient estimator for $\theta$ is defined as \cite[Def. 6.6.6]{lehmann2006theory}
\[
\ARE(\hat{\theta}_n) \triangleq
\lim_{n\rightarrow \infty} \frac{a_n}{n}. 
\]
Note that in the special case where there exists $V \in \mathbb R$ such that
\[
a_n \mathbb E \left(\hat{\theta}_n - \theta \right)^2 = V + o(1),
\]
with $o(1)\to 0$ as $n\to \infty$, then the ARE of $\hat{\theta}_n$ is finite and equals $\sigma^2/V$. \\

@@

\section{Consisten Estimation and Existing Bounds \label{sec:preliminary}}
%On a first impression, it may not be clear whether consistent estimation of the mean is even possible in the adaptive and centralized settings. On the other hand, it may seem as if estimation in these cases is trivial as in the centralized setting (i). 
%
We next settle such skepticism by deriving lower and upper bounds to the relative efficiency under setting (iii).

@@

\subsection{Consistent Estimation}
Fix $\theta_0 \in \mathbb R$ and define the $i$th message by 
\[
B_i = \mathbf 1_{X_i>\theta_0}, 
\]
where $\mathbf 1_A$ is the indicator of the event $A$. We have
\[
p_n \triangleq\frac{1}{n} \sum_{i=1}^n B_i \overset{a.s.}{\rightarrow} F(\theta - \theta_0),  
\]
so that 
\begin{equation}
\label{eq:estimator_naive}
\hat{\theta}_n = \theta_0 + F^{-1}\left( p_n \right)
\end{equation}
is a consistent estimator for $\theta$ in the distributed setting of Figure~\ref{fig:setup}-(iii), where we note that $F(x)$ is invertible over the support of $f(x)$ which is a connected set. Furthermore, the variance of $p_n$ is $F(\theta-\theta_0)F(\theta_0-\theta)$, and hence the delta method implies that $\hat{\theta}_n$ is asymptotically normal with variance 
\begin{equation} 
\frac{1}{
\eta(\theta-\theta_0)} = \frac{F(\theta-\theta_0)F(\theta_0-\theta)}{f^2(\theta-\theta_0)}. 
\end{equation}
In particular, the ARE of $\hat{\theta}_n$ equals $\eta(\theta - \theta_0)\sigma^2$. In other words, for a prescribed accuracy, $\hat{\theta}_n$ of  \eqref{eq:estimator_naive} estimates $\theta$ with sample size that is $\eta(\theta- \theta_0)  \sigma^2$ times the samples size required for the sample mean. \par
%
%
Assumption~\ref{assump:failure_rate} implies that for  all $n$ large enough the ARE of $\hat{\theta}_n$ of \eqref{eq:estimator_naive} is never greater than $\eta(0)$, which is the ARE of the sample median. This ARE is attained only when $\theta_0 = \theta$, although $\theta$ is unknown apriori. Since $\eta(x)$ vanishes as $|x|\rightarrow \infty$, the ARE of $\hat{\theta}_n$ may be very small when $\theta$ is away from $\theta_0$. As an example, when $f(x)$ is a normal density, the ARE of $\hat{\theta}_n$ is $\approx 0.07$ when $\theta_0$ is $2.33$ standard deviations from $\theta$. Therefore, the estimator $\hat{\theta}_n$ has little practical value unless the radius of $\Theta$ is small compared to the standard deviation. \par 
It is suggested that lower variance can be obtained by an estimator of the form \eqref{eq:estimator_naive} if, after observing a batch of the single bit messages, one can update the threshold value $\theta_0$. Such a scheme falls within the adaptive setting of Figure~\ref{fig:setup}-(ii) which we consider in Section~\ref{sec:sequential}. \\
% variance equals to $1/\eta(0)n + o(1)$. We also show in Section~\ref{sec:sequential} that no estimator can attain variance that is asymptotically smaller than $1/\eta(0)n$, so that the ARE of any estimator in settings (ii) and (iii) is at least $\eta(0) \sigma^2$.

\subsection{Multiterminal Source Coding \label{sec:ceo}}
The CEO setting considers the estimation of a sequence $\theta_1,\theta_2\ldots$, where a noisy version of each $\theta_i$ is available at $n$ terminals. At each terminal $i$, an encoder observes the $k$ noisy samples
\[
X_{i,j} = \theta_j + Z_{i,j},\qquad j=1,\ldots,k, \qquad i = 1,\ldots,n,
\]
and transmits $R_i k$ bits to a central estimator  \cite{berger1996ceo}. The central estimator produces an estimates $\hat{\theta}_1,\ldots,\hat{\theta}_k$ with the goal of minimizing the average MSE:
\[
D_{CEO} = \frac{1}{k} \sum_{j=1}^k \mathbb E \left[\left(\theta_j - \hat{\theta}_j \right)^2 \right]. 
\]
Note that any distributed encoding scheme using one-bit per sample can be replicated $k$ times and thus leads to a legitimate encoding and estimation scheme for the CEO problem with $R_1=\ldots=R_n = 1$. It follows that, assuming that $\theta$ is drawn once from the prior $\pi(d\theta)$, our mean estimation problem from one-bit samples under distributed encoding corresponds to the CEO setting with $k=1$ realization of $\theta$ observed under noise at $n$ different locations, and communicated at each location using an encoder sending a single bit. 
%
Consequently, a lower bound on the MSE in estimating $\theta$ in the distributed encoding setting is given by the minimal MSE in the CEO setting as $k \to \infty$. Note that the difference between the CEO setting and ours lays in the privilege of each of the encoders to describe $k$ realizations of $\theta$ using $k$ bits with MSE averaged over these realizations, rather than a single realization using a single bit in ours. 
 \par
When the prior on $\theta$ and the noise corrupting it at each location are Gaussian, the optimal encoding scheme and its asymptotic risk as $k\to \infty$ were fully characterized in \cite{prabhakaran2004rate}. Furthermore, the work of \cite{chen2004upper} provided an expression for MSE attained in the CEO setting under Gaussian priors. Adapting to our setting, this expression provides the following proposition:
\begin{prop} \label{prop:ceo_lower_bound}
Assume that $\Theta = \mathbb R$ and $\pi(\theta) = \mathcal N(0,\sigma_\theta^2)$. Then any estimator $\hat{\theta}_n$ of $\theta$ in the distributed setting satisfies
\begin{equation} \label{eq:ceo_bound}
 n\mathbb E \left( \theta - \theta_n \right)^2 \geq \frac{4\sigma^2}{3} + O(n^{-1}),
\end{equation}
where the expectation is with respect to $\theta$ and $X^n$.
\end{prop}

\begin{proof}
See Appendix. 
\end{proof}

@@

\subsubsection*{Sketch of Proof}
The main idea in the proof is to bound from above the Fisher information of any set of $n$ one-bit messages with respect to $\theta$. Once this bound is achieved, the result follows by using the van-Trees inequality which bounds from below the risk of any estimator of $\theta$ by the inverse of the expected value of the aforementioned Fisher information plus $I_0$. The details are in the Appendix.\\

@@

\subsection{Asymptotically optimal estimator}
Let $\left\{ \gamma_n,\, n\in \mathbb N \right\}$ be a strictly positive sequence. Consider the following estimator $\hat{\theta}_n$ for $\theta$:  
\begin{equation}
\label{eq:sgd_alg}
\theta_n = \theta_{n-1} +  \gamma_n B_n, \quad n = 1,2,\ldots,
\end{equation}
where 
\begin{equation}
B_n = B_n (X_n,\theta_{n-1}) =\sgn (X_n - \theta_{n-1}).
\end{equation}
Define the $n$th step estimation as
\begin{equation} \label{eq:sgd_est}
\hat{\theta}_n =  \frac{1}{n} \sum_{i=1}^n  \theta_i. 
\end{equation}

We have the following results:
\begin{thm} \label{thm:sgd}
Consider the sequence $\left\{\hat{\theta}_n,\, n\in \mathbb N \right\}$ defined by \eqref{eq:sgd_est}. 
\begin{enumerate}
\item[(i)] Assume that $\left\{ \gamma_n,\, n\in \mathbb N \right\}$ satisfies
\begin{equation} \label{eq:conditions1}
\begin{cases}
\frac{\gamma_n - \gamma_{n+1}}{\gamma_n} = o(\gamma_n), &  \\
\sum_{n=1}^\infty \frac{\gamma_n^{(1+\lambda)/2}} {\sqrt{n}} < \infty, & 
\mathrm{for~some~}0< \lambda \leq 1
\end{cases}
\end{equation}
(e.g., $\gamma_n = n^{-\beta}$ for $\beta \in (0,1)$). Then
\[
\sqrt{n} \left( \hat{\theta}_n - \theta \right) \overset{d}{\rightarrow} \mathcal N \left(0,  1/\eta(0) \right).
\]
\item[(ii)] Assume that in addition to \eqref{eq:conditions1}, $\left\{ \gamma_n,\, n\in \mathbb N \right\}$ satisfies
\begin{equation} \label{eq:conditions2}
\begin{cases}
\gamma_n = o(n^{-2/3}), &  \\
\sum_{n=1}^\infty \gamma_n = \infty.  & \\
\end{cases}
\end{equation}
(e.g., $\gamma_n = n^{-\beta}$ with $2/3<\beta<1$). Then
\[
\lim_{n\rightarrow \infty} n\mathbb E \left[ \left(\theta-\hat{\theta}_n \right)^2 \right] = \frac{1}{\eta(0)}. 
\]
\end{enumerate}
\end{thm}

\subsubsection*{Proof}
The asymptotic behavior of \eqref{eq:sgd_est} follows from \cite[Thm. 4]{polyak1992acceleration} and \cite[Thm. 2]{polyak1990new}. The details are in the Appendix. \\

Theorem~\ref{thm:sgd} implies that the estimator $\hat{\theta}_n$, defined by \eqref{eq:sgd_est} and \eqref{eq:sgd_alg}, attains the maximal ARE as established by Theorem~\ref{thm:adpative_lower_bound}. The update step \eqref{eq:sgd_alg} can be seen as a gradient descent step for the function $x\to |x|$ at the point $x=X_n - \theta_{n-1}$. Consequently, the procedure above is known as averaged stochastic gradient descent for minimizing $x \to |x|$ given the data $X_1,\ldots,X_n$. The minimal value of this optimization is the sample median, and Theorem~\ref{thm:sgd} provides conditions for the sequence of gradient steps so that the algorithm converges to this minimum. 
 \par
%Note that $\theta_0$ is not explicitly defined in equation \eqref{eq:sgd_est}. A reasonable initialization for $\theta_0$ is $\theta_0 = \mathbb E [\theta]$, although Theorem~\ref{thm:sgd} implies that the asymptotic behavior of the estimator is indifferent to this initialization. Thus, the optimal efficiency is attained regardless of the prior distribution on $\theta$ or the radius of the parameter space $\Theta$.  \\
In the encoding and estimating procedure \eqref{eq:sgd_alg} and \eqref{eq:sgd_est}, each one-bit message $B_n$ depends on its private sample as well as the current gradient descent estimate $\theta_{n-1}$. In this sense, each encoder in this algorithm interacts with previous one by using the current estimate.
%
As we explain next, 
it is possible to obtain the optimal efficiency of $\eta(0)\sigma^2$ with only one round of interactions among the encoders.

@@

\node[below right = 0.75 and 1.5 of source3] (sourceB) {$X_{n_1 +1}$} ;
\node[int1, right of = sourceB, node distance = 1.7cm] (enc1B) {Enc$(\hat{\theta}_1)$};  
\draw[->,line width = 2pt] (sourceB) -- (enc1B);

\node[below of = sourceB, node distance = 1.7cm] (source3B) {$X_n$};
\node[int1, right of = source3B, node distance = 1.7cm] (enc3B) {Enc$(\hat{\theta}_1)$};  
\draw[->,line width = 2pt] (source3B) -- (enc3B); 
\node[below of = sourceB, node distance = 0.4cm] {$\vdots$};

@@

\begin{proof}
See Appendix. 
\end{proof}

@@

\caption{\label{fig:minimax_support}
Optimal threshold density $\lambda^\star(dt)$ (blue) that maximizes the ARE for $f(x) = \mathcal N(\theta,\sigma^2)$ and $\theta \in \Theta = [-T,T]$.
%
The continuous curve (red) represents the reciprocal of the asymptotic risk for at a fixed $\theta \in \Theta$ under the optimal density, so the minimax risk is the inverse of its minimal value. The dashed curve (green) is the reciprocal of the asymptotic risk for a fixed $\theta$ under a uniform distribution of threshold values over $\Theta$, hence its minimal value is the inverse of \eqref{eq:uniform_risk}. }
\end{center}
\end{figure}

@@

\addplot[color = black!35!green, smooth, dashed, line width = 1pt] plot table [x = x, y = z, col sep=comma] {./Figs/minmax_ARE_b0.5.csv};
\addlegendentry{uniform threshold desnity};
\end{axis}
\end{tikzpicture}
\caption{\label{fig:minimax_ARE} Minimax ARE versus $\sigma/T$ for $f(x) = \mathcal N(\theta,\sigma^2)$  and $\theta \in \Theta = [-T,T]$. The dashed curve (green) is the ARE under a uniform threshold density over $\Theta$ given by $K_{\unif}\sigma^2$, where $\kappa_{\unif}$ is given by \eqref{eq:uniform_risk}. }.
\end{center}
\end{figure}

@@

 We consider now the problem of minimizing the asymptotic Bayes risk $
R_{\pi,\lambda} \triangleq \mathbb E  K^{-1}(\theta)$ over all probability measures $\lambda(dt)$ with support in $\mathbb R$. This optimization problem can be written as follows:
\begin{align}
\label{eq:cvx_average}
\begin{split}
\mathrm{minimize} \quad & R_{\pi,\lambda} =  \int \frac{\pi(d\theta)}{ \int \eta \left( t-\theta\right) \lambda(dt)}. \\ 
\mathrm{subject~to} \quad & \lambda(dt)\geq 0,\quad \int \lambda(dt) =1. 
\end{split}
\end{align}
We denote by $R^\star_{\pi}$ the minimal value of the objective function in \eqref{eq:cvx_average}. Since the function $x \rightarrow 1/x$ is convex for positive values, \eqref{eq:cvx_average} defines a convex optimization problem in $\lambda$ whose solution depends on the prior $\pi$. The solution to this problem is approximated by considering $\lambda$ and $\pi$ over a discrete set. On Figure~\ref{fig:opt_density} we illustrate the solution to \eqref{eq:cvx_average} for the case of where $f(x)$ is the normal distributon ($P_X = \mathcal N(\theta, \sigma^2)$) and the prior on $\theta$ is uniform over $\Theta = [-b,b]$. Figure~\ref{fig:dist_bound_uniform} illustrates the corresponding Bayes risk. 
\par
\begin{figure}
\begin{center}

>>>>

\documentclass[letterpaper, 11pt]{IEEEtran}      % Use this line for a4 paper

@@

\usepackage[hidelinks]{hyperref}
%\tikzexternalize[mode=list and make]
\usepackage{algorithmicx}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage{pstool}
\usepackage[latin1]{inputenc}
\usetikzlibrary{arrows,shapes}
\usepackage{xifthen}
\usepackage{epic}
%\usepackage{caption}
\usepackage{epstopdf}

@@

\newcommand{\mmse}{\mathsf{mmse}}
\newcommand{\ex}[1]{\ensuremath{\mathbb{E}\left[ #1\right]}}

\newcommand{\Var}[1]{\ensuremath{\mathrm{Var} \left[ #1\right]}}

\newcommand{\unif}{\mathsf{unif}}
\newcommand{\card}{\mathrm{card}}
\newcommand{\ARE}{\mathsf{ARE}}
\newcommand{\supp}{\mathrm{supp} }
\renewcommand\vec[1]{\ensuremath\boldsymbol{#1}}
\newenvironment{proof}{\paragraph*{Proof}}{\hfill$\square$ \newline}
\newcommand{\sgn}{\mathrm{sgn} }
\newcommand{\Prob}{\mathbb{P} }
\newcommand{\argmax}{\mathrm{argmax}}
\newcommand*{\QEDA}{\hfill\ensuremath{\square}}
\newcommand{\CEO}{\mathsf{CEO}}
\newcommand{\Ncal}{\mathcal{N}}

@@

%\node[above of = source3, node distance = 1cm] (dist) {$X_i \sim {\Ncal} \left(\theta, \sigma^2 \right)$};
\node[below of = source2, node distance = 0.5cm] {$\vdots$};

@@

We give particular attention to the ARE of estimators with respect to an asymptotically normal efficient estimator that is not subject to the bit constraint. Specifically, let $\{a_n,\,n\in \mathbb N\}$ be a sequence such that 
\[
\sqrt{a_n}\left(\hat{\theta}_n - \theta\right) \overset{d}{\longrightarrow} \Ncal(0, \sigma^2).
\]
Then the ARE of $\hat{\theta}_n$ with respect to an unconstrained efficient estimator for $\theta$ is defined as \cite[Def. 6.6.6]{lehmann2006theory}
\[
\ARE(\hat{\theta}_n) \triangleq
\lim_{n\rightarrow \infty} \frac{a_n}{n}. 
\]
Note that in the special case where there exists $V \in \mathbb R$ such that
\[
a_n \mathbb E \left(\hat{\theta}_n - \theta \right)^2 = V + o(1),
\]
with $o(1)\to 0$ as $n\to \infty$, then the ARE of $\hat{\theta}_n$ is finite and equals $\sigma^2/V$. \\

@@

\section{Consistent Estimation and Off-the-shelf Bounds \label{sec:preliminary}}
%On a first impression, it may not be clear whether consistent estimation of the mean is even possible in the adaptive and centralized settings. On the other hand, it may seem as if estimation in these cases is trivial as in the centralized setting (i). 
%
We next settle such skepticism by deriving lower and upper bounds to the relative efficiency under setting (iii).

@@

\subsection{Consistent Estimation}
Fix $\theta_0 \in \mathbb R$ and define the $i$th message by 
\[
B_i = \mathbf 1_{X_i>\theta_0}, 
\]
where $\mathbf 1_A$ is the indicator of the event $A$. We have
\[
P_n \triangleq\frac{1}{n} \sum_{i=1}^n B_i \overset{a.s.}{\rightarrow} F(\theta - \theta_0),  
\]
so that 
\begin{equation}
\label{eq:estimator_naive}
\hat{\theta}_n = \theta_0 + F^{-1}\left( P_n \right)
\end{equation}
is a consistent estimator for $\theta$ in the distributed setting of Figure~\ref{fig:setup}-(iii), where we note that $F(x)$ is invertible over the support of $f(x)$ which is a connected set due to log-concavity. Furthermore, the variance of $P_n$ is $F(\theta-\theta_0)F(\theta_0-\theta)$, and hence the delta method implies that $\hat{\theta}_n$ is asymptotically normal with variance 
\begin{equation} 
 \frac{F(\theta-\theta_0)F(\theta_0-\theta)}{f^2(\theta-\theta_0)} = \frac{1}{
\eta(\theta-\theta_0)}.
\end{equation}
In particular, the ARE of $\hat{\theta}_n$ equals $\eta(\theta - \theta_0)\sigma^2$. In other words, for a prescribed accuracy, $\hat{\theta}_n$ of  \eqref{eq:estimator_naive} estimates $\theta$ with sample size that is $\eta(\theta- \theta_0)  \sigma^2$ times the samples size required for the sample mean. \par
%
%
Assumption~\ref{assump:failure_rate} implies that for  all $n$ large enough the ARE of $\hat{\theta}_n$ of \eqref{eq:estimator_naive} is never greater than $\eta(0)$, which is the ARE of the sample median. This ARE is attained only when $\theta_0 = \theta$, whereas $\theta$ is apriori unknown. Since $\eta(x)$ vanishes as $|x|\rightarrow \infty$, the ARE of $\hat{\theta}_n$ may be very small when $\theta$ is away from $\theta_0$. As an example, when $f(x)$ is a normal density, the ARE of $\hat{\theta}_n$ is less than $\approx 0.14$ when $\theta_0$ is $2$ standard deviations from $\theta$. Therefore, the estimator $\hat{\theta}_n$ has little practical value unless the radius of $\Theta$ is small compared to the standard deviation. \par 
It is suggested that we can attain lower variance in estimation by updating the threshold value $\theta_0$ in \eqref{eq:estimator_naive} after observing one or a batch of the single bit messages. Schemes with such adaptation fall within the adaptive setting of Figure~\ref{fig:setup}-(ii) which we consider in Section~\ref{sec:sequential}. \\

\subsection{Multiterminal Source Coding \label{sec:ceo}}
The CEO setting considers the estimation of a sequence $\theta_1,\theta_2\ldots$, where a noisy version of each $\theta_i$ is available at $n$ terminals. At each terminal $i$, an encoder observes the $k$ noisy samples
\[
X_{i,j} = \theta_j + Z_{i,j},\qquad j=1,\ldots,k, \qquad i = 1,\ldots,n,
\]
and transmits $R_i k$ bits to a central estimator  \cite{berger1996ceo}. The central estimator produces an estimates $\hat{\theta}_1,\ldots,\hat{\theta}_k$ with the goal of minimizing the average MSE:
\[
MSE_{\CEO} = \frac{1}{k} \sum_{j=1}^k \mathbb E \left[\left(\theta_j - \hat{\theta}_j \right)^2 \right]. 
\]
Note that any distributed encoding scheme using one-bit per sample can be replicated $k$ times and thus leads to a legitimate encoding and estimation scheme for the CEO problem with $R_1=\ldots=R_n = 1$. It follows that, assuming that $\theta$ is drawn once from the prior $\pi(d\theta)$, our mean estimation problem from one-bit samples under distributed encoding corresponds to the CEO setting with $k=1$ realization of $\theta$ observed under noise at $n$ different locations, and communicated at each location using an encoder sending a single bit. 
%
Consequently, a lower bound on the MSE in estimating $\theta$ in the distributed encoding setting is given by the minimal MSE in the CEO setting as $k \to \infty$. Note that the difference between the CEO setting and ours lays in the privilege of each of the encoders to describe $k$ realizations of $\theta$ using $k$ bits with MSE averaged over these realizations, rather than a single realization using a single bit in ours. 
 \par
When the prior on $\theta$ and the noise corrupting it at each location are Gaussian, the optimal encoding scheme and its asymptotic risk as $k\to \infty$ were fully characterized in \cite{prabhakaran2004rate}. Furthermore, the work of \cite{chen2004upper} provided an expression for MSE attained in the CEO setting under Gaussian priors. Adapting to our setting, this expression provides the following proposition:
\begin{prop} \label{prop:ceo_lower_bound}
Assume that $\Theta = \mathbb R$ and $\pi(\theta) = \Ncal(0,\sigma_\theta^2)$. Then any estimator $\hat{\theta}_n$ of $\theta$ in the distributed setting satisfies
\begin{equation} \label{eq:ceo_bound}
 n\mathbb E \left( \theta - \theta_n \right)^2 \geq \frac{4}{3} \sigma^2 + O(n^{-1}),
\end{equation}
where the expectation is with respect to $\theta$ and $X^n$.
\end{prop}

\begin{proof}
See Appendix~\ref{app:proof:prop_ceo_lower_bound}.
\end{proof}

@@

\subsubsection*{Sketch of Proof}
The main idea in the proof is to bound from above the Fisher information of any set of $n$ one-bit messages with respect to $\theta$. Once this bound is achieved, the result follows by using the van-Trees inequality which bounds from below the risk of any estimator of $\theta$ by the inverse of the expected value of the aforementioned Fisher information plus $I_0$. The details are in Appendix~\ref{proof:thm:adpative_lower_bound}.\\

@@

\subsection{Asymptotically optimal estimator}
Let $\left\{ \gamma_n,\, n\in \mathbb N \right\}$ be a strictly positive sequence. Consider the following estimator $\hat{\theta}_n$ for $\theta$:  
\begin{equation}
\label{eq:sgd_alg}
\theta_n = \theta_{n-1} +  \gamma_n B_n, \quad n = 1,2,\ldots,
\end{equation}
where 
\begin{equation}
B_n = B_n (X_n,\theta_{n-1}) =\sgn (X_n - \theta_{n-1}).
\end{equation}
Define the $n$th step estimation as
\begin{equation} \label{eq:sgd_est}
\bar{\theta}_n =  \frac{1}{n} \sum_{i=1}^n  \theta_i. 
\end{equation}

We have the following results:
\begin{thm} \label{thm:sgd}
Consider the sequence $\left\{\bar{\theta}_n,\, n\in \mathbb N \right\}$ defined by \eqref{eq:sgd_est}. 
\begin{enumerate}
\item[(i)] Assume that $\left\{ \gamma_n,\, n\in \mathbb N \right\}$ satisfies
\begin{equation} \label{eq:conditions1}
\begin{cases}
\frac{\gamma_n - \gamma_{n+1}}{\gamma_n} = o(\gamma_n), &  \\
\sum_{n=1}^\infty \frac{\gamma_n^{(1+\lambda)/2}} {\sqrt{n}} < \infty, & 
\mathrm{for~some~}0< \lambda \leq 1
\end{cases}
\end{equation}
(e.g., $\gamma_n = n^{-\beta}$ for $\beta \in (0,1)$). Then
\[
\sqrt{n}\left( \bar{\theta}_n - \theta\right) \overset{d}{\to} \Ncal\left(0,\frac{1}{\eta(0)}\right).
\]
\item [(ii)] Assume in addition that $f(x)$ is continuously differentiable with finite Fisher information for location
\[
I_\theta = \int_{\mathbb R} \left( \frac{f'(x)}{f(x)}  \right)^2 f(x) dx < \infty.
\]
%(or that $f(x-\theta)$ is differentiable in quadratic mean).  
For $\theta \in \Theta$, $h\in \mathbb R$ and $n$ large enough such that $\theta+h/\sqrt{n} \in \Theta$, let $P^n_{\theta,h,n}$ be the measure with density $f(x-\theta - h/\sqrt{n})$. Then 
\begin{align}
\label{eq:sgd_part2}
\sqrt{n}\left( \bar{\theta}_n - \theta\right) \overset{d}{\to} \Ncal\left( h,\frac{1}{\eta(0)}\right),
\end{align}
under $P^n_{\theta,h,n}$. 
\item[(iii)] Assume that, in addition to \eqref{eq:conditions1}, $\left\{ \gamma_n,\, n\in \mathbb N \right\}$ satisfies
\begin{equation} \label{eq:conditions2}
\begin{cases}
\gamma_n = o(n^{-2/3}), &  \\
\sum_{n=1}^\infty \gamma_n = \infty.  & \\
\end{cases}
\end{equation}
(e.g., $\gamma_n = n^{-\beta}$ with $\beta \in (2/3,1)$). Then
\[
\lim_{n\to \infty} n \ex{\left( \bar{\theta}_n - \theta \right)^2} = \frac{1}{\eta(0)}. 
\]
\end{enumerate}
\end{thm}

\subsubsection*{Proof}
See Appendix~\ref{proof:thm:sgd}. \\

Theorem~\ref{thm:sgd} implies that the estimator $\hat{\theta}_n$, defined by \eqref{eq:sgd_est} and \eqref{eq:sgd_alg}, attains the maximal ARE as established by Theorem~\ref{thm:adpative_lower_bound}. Furthermore, it is locally minimax in the sense that it attains the minimal MSE for all local alternatives of $\theta$. 
%
The update step \eqref{eq:sgd_alg} can be seen as a gradient descent step for the function $x\to |x|$ at the point $x=X_n - \theta_{n-1}$. Consequently, the procedure above is known as averaged stochastic gradient descent for minimizing $x \to |x|$ given the data $X_1,\ldots,X_n$. The minimal value of this optimization is the sample median, and Theorem~\ref{thm:sgd} provides conditions for the sequence of gradient steps so that the algorithm converges to this minimum. 
 \par
%Note that $\theta_0$ is not explicitly defined in equation \eqref{eq:sgd_est}. A reasonable initialization for $\theta_0$ is $\theta_0 = \mathbb E [\theta]$, although Theorem~\ref{thm:sgd} implies that the asymptotic behavior of the estimator is indifferent to this initialization. Thus, the optimal efficiency is attained regardless of the prior distribution on $\theta$ or the radius of the parameter space $\Theta$.  \\
In the encoding and estimating procedure \eqref{eq:sgd_alg} and \eqref{eq:sgd_est}, each one-bit message $B_n$ depends on its private sample as well as the current gradient descent estimate $\theta_{n-1}$. In this sense, each encoder in this algorithm interacts with previous one by using the current estimate.
%
As we explain next, 
it is possible to obtain the optimal efficiency of $\eta(0)\sigma^2$ with only one round of interactions among the encoders.

@@

\node[below right = 0.75 and 1.5 of source3] (sourceB) {$X_{n_1 +1}$} ;
\node[int1, right of = sourceB, node distance = 1.7cm] (enc1B) {Enc$(\hat{\theta}_{n_1})$};  
\draw[->,line width = 2pt] (sourceB) -- (enc1B);

\node[below of = sourceB, node distance = 1.7cm] (source3B) {$X_n$};
\node[int1, right of = source3B, node distance = 1.7cm] (enc3B) {Enc$(\hat{\theta}_{n_1})$};  
\draw[->,line width = 2pt] (source3B) -- (enc3B); 
\node[below of = sourceB, node distance = 0.4cm] {$\vdots$};

@@

\begin{proof}
See Appendix~\ref{proof:non_existence}
\end{proof}

@@

\caption{\label{fig:minimax_support}
Optimal threshold density $\lambda^\star(dt)$ (blue) that maximizes the ARE for $f(x) = \Ncal(\theta,\sigma^2)$ and $\theta \in \Theta = [-T,T]$.
%
The continuous curve (red) represents the reciprocal of the asymptotic risk for at a fixed $\theta \in \Theta$ under the optimal density, so the minimax risk is the inverse of its minimal value. The dashed curve (green) is the reciprocal of the asymptotic risk for a fixed $\theta$ under a uniform distribution of threshold values over $\Theta$, hence its minimal value is the inverse of \eqref{eq:uniform_risk}. }
\end{center}
\end{figure}

@@

\addplot[color = black!35!green, smooth, dashed, line width = 1pt] plot table [x = x, y = z, col sep=comma] {./Figs/minmax_ARE_b0.5.csv};
\addlegendentry{uniform threshold desnity};
\end{axis}
\end{tikzpicture}
\caption{\label{fig:minimax_ARE} Minimax ARE versus $\sigma/T$ for $f(x) = \Ncal(\theta,\sigma^2)$  and $\theta \in \Theta = [-T,T]$. The dashed curve (green) is the ARE under a uniform threshold density over $\Theta$ given by $K_{\unif}\sigma^2$, where $\kappa_{\unif}$ is given by \eqref{eq:uniform_risk}. }.
\end{center}
\end{figure}

@@

 We consider now the problem of minimizing the asymptotic Bayes risk $
R_{\pi,\lambda} \triangleq \mathbb E  K^{-1}(\theta)$ over all probability measures $\lambda(dt)$ with support in $\mathbb R$. This optimization problem can be written as follows:
\begin{align}
\label{eq:cvx_average}
\begin{split}
\mathrm{minimize} \quad & R_{\pi,\lambda} =  \int \frac{\pi(d\theta)}{ \int \eta \left( t-\theta\right) \lambda(dt)}. \\ 
\mathrm{subject~to} \quad & \lambda(dt)\geq 0,\quad \int \lambda(dt) =1. 
\end{split}
\end{align}
We denote by $R^\star_{\pi}$ the minimal value of the objective function in \eqref{eq:cvx_average}. Since the function $x \rightarrow 1/x$ is convex for positive values, \eqref{eq:cvx_average} defines a convex optimization problem in $\lambda$ whose solution depends on the prior $\pi$. The solution to this problem is approximated by considering $\lambda$ and $\pi$ over a discrete set. On Figure~\ref{fig:opt_density} we illustrate the solution to \eqref{eq:cvx_average} for the case of where $f(x)$ is the normal distributon ($P_X = \Ncal(\theta, \sigma^2)$) and the prior on $\theta$ is uniform over $\Theta = [-b,b]$. Figure~\ref{fig:dist_bound_uniform} illustrates the corresponding Bayes risk. 
\par
\begin{figure}
\begin{center}

\begin{tikzpicture}[scale = 1]
\begin{axis}[
width=8cm, height=6cm,
xmin = 0, xmax=2, 
restrict y to domain = 0:100,
ymin = 0,
ymax = 10,
samples=1, 
xlabel= {$\sigma/\sigma_\theta$},
%xtick={-0.5,0,0.5},
%xticklabels={-$$, 0, $b$},
ytick={0,1.5708,5},
yticklabels={0,$\pi/2$,5},
ylabel = {$R_{\pi}^\star/\sigma^2$},
line width=1.0pt,
mark size=1.5pt,
ymajorgrids,
xmajorgrids,
legend style= {at={(1,1)},anchor=north east,draw=black,fill=white,align=left}
]
\addplot[color = red, smooth] plot table [x = x, y = Runif, col sep=comma] {./Figs/unif_Bayes_Risk.csv};
%\addlegendentry{$\frac{\sigma^2}{\sigma_\theta^2} = 1$};
\addplot[color = red, smooth, dashed, line width = 0.5pt] plot table [x = x, y = Rbound, col sep=comma] {./Figs/unif_Bayes_Risk.csv};
\addplot[color = black, dashed] {3.14159 / 2};
\end{axis}
\end{tikzpicture}
\caption{Asymptotic Bayes risk $R_{\pi}^\star$ in estimating the mean of a normal distribution ($P_X = \Ncal(\theta, \sigma^2)$under an optimal threshold distribution $\lambda^\star$ for $\pi$ the uniform distribution over $\Theta = [-0.5,0.5]$. 
%
%versus $\sigma/\sigma_\theta = 2\sqrt{3}\sigma$.
The distribution $\lambda^\star$ is the minimizer of \eqref{eq:cvx_average}. It is illustrated for various cases in Fig.~\ref{fig:opt_density}. The dashed curve represents the upper bound \eqref{eq:upper_bound}. 
\label{fig:dist_bound_uniform}  }
\end{center}
\end{figure}
